{DEM}-{SPH} study of particles dispersion in fluid

### 2024-05-30 manuscript/paper.tex
Correct the draft

\documentclass[preprint,12pt]{elsarticle}
% \documentclass[draft,12pt]{elsarticle}

@@

  \title{Study of mixing behaviour of spherical particles in fluid using stirrer}
  \author[XXX]{Dinesh Adepu\corref{cor1}}
  \ead{d.dinesh@surrey.ac.uk}
  \author[University of Surrey]{Chuan Yu Wu}
  \ead{XXX}
\address[xxx]{xxx}

@@

\begin{abstract}

\end{abstract}

@@

Particle dispersion is a fundamental physical phenomenon encountered in
various industries where powder and fluid interactions are pivotal.  Petroleum
industry \cite{wang2023developments}, The transport of bodies in internal
systems \cite{Dai2021}, debris flow \cite{Qingyun2022}, the food processing
industry \cite{Karunasena2014}, and ice-sea modeling \cite{Mintu2018} are a
few areas to mention.  Particularly, processes involving the mixing of powder
with fluid using a stirrer are commonplace \cite{li2022study}. While
experimental investigation of such processes is often impractical due to their
highly nonlinear nature, numerical methods offer a viable alternative. These
systems are studied as part of two-way coupling models. The two-way coupling
phenomena are nonlinear, and an analytical study is not feasible. A numerical
study is preferable while handling such a phenomenon. A mesh-based or meshless
technique can be utilized to study rigid fluid coupling (RFC) numerically.
\Cref{fig:gloabl_body_frame_rb} illustrates a sample particle dispersion
problem studied in the current work.
\begin{figure}[!htpb]
  \centering
  \includegraphics[width=0.4\textwidth]{images/hs_tank_fluid_with_particles}
  \caption{An illustrative figure showcasing the dispersion of particles within a free surface tank.}
  \label{intro:schematic}
\end{figure}

The combination of Discrete Element Method (DEM) for solid particle
interactions and Computational Fluid Dynamics (CFD) for fluid modeling
is a classic approach in handling particulate flow problem. In
mesh-based CFD-DEM modeling, lattice Boltzmann method (LBM)
\cite{xiong2014lbm} and finite volume method (FVM) \cite{kloss2012models} are
utilized to handle the fluid dynamics. Similarly, in meshless schemes like
Smoothed Particle Hydrodynamics (SPH) \cite{peng2021fully},
Moving Particle Semi-implicit (MPS), and Particle Finite Element
Method (PFEM) \cite{li2019modeling, franci2020pfem}, handles the fluid
dynamics.  While handling the interaction between the fluid and solid
particles, the interaction can be categorized into a resolved and unresolved
coupling. In unresolved coupling, simplified models compute forces acting on
solid particles within fluid flow, while fully resolved coupling computes
forces comprehensively. Both strategies are applicable in both meshless
\cite{cleary2015prediction, trujillo2020smooth} and mesh-based
\cite{van2008numerical, ma2022review} CFD-DEM techniques. Research such as
\cite{brosh2014accelerating} and \cite{peng2021fully} explore wide particle
size distributions with CFD-DEM and SPH-DEM respectively.  Regarding
non-Newtonian fluid modeling in mesh-based schemes, works like
\cite{li2018dam} are notable and in meshless schemes, \cite{peng2021fully} is
notable. \citet{zhu2008discrete} and \citet{ma2022review}
delve into investigations on spherical and non-spherical particles
respectively within CFD-DEM frameworks.  Meshless CFD-DEM, though fruitful,
encounters challenges in handling problems with free surfaces and becomes
computationally demanding with a large number of solid particles, especially
in fully resolved coupling. In contrast, SPH offers advantages in such
scenarios due to its inherent capability in handling free surfaces and
adeptness in modeling large mesh deformation problems, facilitating the
incorporation of novel physics.

The SPH method is a meshless numerical method originally proposed by
\citet{gingold1977smoothed} and \citet{lucy1977numerical} to model
astrophysics problems. It has been extensively applied to simulate problems
involving fluids, structural dynamics, fluid-structure interaction, granular
physics, non-Newtonian fluid flows \cite{peng2021fully} among other areas
\cite{monaghan2012smoothed}.  Various kind of particulate flows, involving,
simple spherical particles mixing using unresolved coupling
\cite{cleary2015prediction, markauskas2019coupled} and with debris
flow \cite{canelas2016sph} with fully resolved coupling approach has been
solved using SPH-DEM framework.  Handling particles of arbitrary shape in
fluid flow is addressed in various works, including those by
\citet{peng2021fully, canelas2016sph, amicarelli2015smoothed}.  Several
variants of SPH, such as WCSPH \cite{peng2021fully, cleary2015}, and ISPH
\cite{asai2021fluid}, are utilized for modeling fluid flow with particulate
flows. The coupling between fluid and rigid bodies in the fully resolved
category employs techniques such as the fixed ghost particle technique
\cite{canelas2016sph, asai2021fluid}, single layers of dummy SPH particles
\cite{peng2021fully}, and simple repulsive forces
\cite{monaghan2009sph}. Coupling between SPH fluid particles and solid
structures discretized as vertex edges is proposed by
\cite{park2023new}. Particle settling with variable sizes is studied by
\citet{zou2022study}.  Elastic behavior of spherical particles and dispersion
studies are conducted by \citet{ng2021numerical}.  Interaction between rigid
bodies of arbitrary sizes is addressed using various techniques, such as those
discussed in papers by \citet{asai2021fluid}, citet{peng2021fully},
\citet{canelas2016sph}.

In the current work we cosider the particle dispersion within a large-scale
under stirrer mixing phenomenon. We employ Smoothed Particle Hydrodynamics
(SPH) to model fluid dynamics and the Discrete Element Method (DEM) to handle
the dynamics and contact interactions among spherical particles. The
interaction between solids and fluids is managed using a fixed ghost particle
approach, ensuring a fully resolved simulation. We validate the fluid solver
independently using Poiseuille flow and the DEM solver separately using
fundamental benchmarks such as particle-particle and particle-wall
impacts. Subsequently, the fully resolved coupled SPH-DEM solver is validated
using three benchmarks: a single particle entering, two particles settling,
and a forced wedge entry into a steady hydrostatic tank. Upon confirming the
fidelity of the solver, we proceed with analyzing the mixing behavior of
various spherical particles influenced by a stirrer in a water tank. We
investigate scenarios involving particles of different densities, stirrer
velocities, and variable sizes as part of our case study. The development work
is conducted using PySPH\cite{ramachandran2021pysph}, an open-source code
available at \url{https://github.com/pypr/pysph} . To ensure reproducibility,
we utilize the automan package \cite{ramachandran2018automan} to automate all
results generated in the current manuscript, aiming for a reproducible
research approach.

@@

\FloatBarrier%
\section{Fluid modeling}
\label{sec:fluid-modeling}
In the current work, we follow a weakly-compressible SPH approach to model the
fluid. Smoothed particle hydrodynamics is initially developed by Lucy
\cite{lucy1977numerical} and Gingold and Monaghan \cite{gingold1977smoothed}
to model astrophysical problems. The continuum in SPH is modeled using
particles, which has physical properties such as mass, velocity, and the
particles interact based on the governing equations using a Guassian-like
kernel \cite{SPH_papers}.

\FloatBarrier%
\subsection{Fluid governing equations}
\label{sec:fluid--governing-equations}
Equations governing the fluid motion are, conservation of mass
\begin{equation}
  \label{eq:ce}
  \frac{d \rho}{d t} = - \rho \; \frac{\partial u_i}{\partial x_i},
\end{equation}
and conservation of linear momentum,
\begin{equation}
  \label{eq:me}
  \frac{d u_i}{d t} = \frac{1}{\rho} \; \frac{\partial \sigma_{ij}}{\partial x_j}
  + g_i,
\end{equation}
where $\rho$ is the density, $u_i$ is the $i$\textsuperscript{th} component of
the velocity field, $x_j$ is the $j$\textsuperscript{th} component of the
position vector, $g_i$ is the component of body force per unit mass and
$\sigma_{ij}$ is stress tensor.

The stress tensor is split into a pressure and viscous parts,
\begin{equation}
  \label{eq:fluid-stress-decomposition}
  \sigma_{ij} = - p \delta_{ij} + 2 \eta \frac{\partial u_i}{\partial x_j}
\end{equation}
where $\eta$ is the kinematic viscosity of the fluid, $p$ is the pressure,
$\delta_{ij}$ is the Kronecker delta function.

\FloatBarrier%
\subsection{Discretized fluid governing equations}
\label{sec:sph--governing-equations}
The governing equations involve function, derivative and divergence
operators. In SPH, these operators are approximated based on the positions,
mass and the kernel values of the discretized particles. Assume that the
domain is discretized into N particles with mass $m_i$ and volume
$\Delta v_i$, We have
\begin{equation}
  \label{eq:mass_repr}
  m_j = \Delta v_i \> \rho_i.
\end{equation}
Based on this discretization, the discrete function approximation is given as,
\begin{equation}
  \label{eq:discrete_form}
  A_i = \sum_{j \in \text{Neigh}(i)}\> \frac{m_j}{\rho_j} A_j\> W(\ten{x}_i - \ten{x}_j, h),
\end{equation}
where $A(\boldsymbol{x}_i) = A_i$ is the value of the field property of
particle i, similarly for $A_j$. Here, $\text{Neigh}(i)$ is the neighbours of
particle $i$.  A symmetric derivative approximation~\cite{Violeau16} is given
as,
\begin{equation}
  \nabla A(\ten{x}_i) = \rho_i \sum_{j} m_j \left(\frac{A_i}{\rho_i^2} + \frac{A_j}{\rho_j^2}\right) \nabla W_{ij}.
\end{equation}
A symmetric gradient operator ensures equal and opposite forces acting on the
particles.

@@

% \subsection{Discretized fluid governing equations}
The SPH discretization of the continuity
equation~\cref{eq:sph-discretization-continuity} and the momentum equation
~\cref{eq:sph-momentum-fluid} respectively are,
\begin{equation}
  \label{eq:sph-discretization-continuity}
  \frac{{d}\rho_a}{dt} = \sum_{b} \; \frac{m_b}{\rho_{b}} \;
  \rho_{a} \; {\ten{u}}_{ab} \; \cdot \nabla_{a} W_{ab},
\end{equation}

%
Similarly, the discretized momentum equation for fluids is written as,
\begin{multline}
  \label{eq:sph-momentum-fluid}
  \frac{{d}\ten{u}_{a}}{dt} = - \sum_{b} m_b
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg)
  \nabla_{a} W_{ab}
 \;+\;
  \sum_{b} m_b \frac{4 \eta \nabla W_{ab}\cdot
    \ten{r}_{ab}}{(\rho_a + \rho_b) (r_{ab}^2 + 0.01 h_{ab}^2)} \ten{u}_{ab}  \;+\;
  \ten{g}_{a},
\end{multline}
where $\ten{I}$ is the identity matrix, $\eta$ is the kinematic viscosity of the
fluid and \cite{morris1997modeling} formulation is used to discretize the
viscosity term.

@@

The ghost particle approach of \cite{Adami2012} is used to model the
boundaries. We use three layers of ghost particles to model the solid wall.
The properties of the solid wall are interpolated from the fluid particles.

When the viscous force is computed, the no slip boundary condition is used,
where the velocity on the boundary set as,
\begin{equation}
  \label{eq:no-slip-bc-u}
  \ten{u}_{\text{Ga}} = 2 \ten{u}_{\text{p}} - \ten{\hat{u}}_{\text{a}}.
\end{equation}
The projected velocity $\ten{\hat{u}}_{\text{a}}$ on the ghost particles is
computed using,
\begin{equation}
  \label{eq:v-ghost}
  \ten{\hat{u}}_a = \frac{\sum_b\ten{u}_b W_{ab}}{\sum_b W_{ab}},
\end{equation}
where $\ten{u}_b$, is the momentum velocity
of the fluid and $W_{ab}$ is the kernel value between the fluid
particle and the ghost particle.

@@

\FloatBarrier%
\subsection{Contact resolution}
\label{sec:dem}

@@

We resolve the contact among the spherical particles using discrete element
method. In the current work we have utilized a non-linear contact force
model. In DEM, the force acting on a particle $a$ due to the interaction with
the particle $b$ is resolved into a normal and tangential component. The
normal force component is utilised to make sure that the particle do not
penetrate into each other, while the tangential component is used to model the
friction between the interacting particles.  The normal force
($\teng{F}_a^{n}$) on particle $a$ due to the interaction with the particles
$b$ is given by a non-linear, Hetzian model \citet{brilliantov1996model},
\begin{equation}
  \label{eq:contact-algorithm-normal}
  \ten{F}_a^n = k_r \delta_{n} \ten{n}.
\end{equation}
Here, the overlap $\delta_{n}$ is computed using
\begin{equation}
  \label{eq:cf-overlap}
  \delta_{n} = R_{a} + R_{b} - r_{ab},
\end{equation}
$k_r$ is the normal spring stiffness coefficient, which is computed using the
material properties of the bodies in contact \cite{golshan2023lethe}.

\subsection{Tangential force computation}
\label{sec:tangential-force-computation}
To handle the frictional contact, we associate a tangential spring attached to
particle $a$ and particle $b$ to compute the tangential force, which initially has
a magnitude of zero ($|\Delta \textit{\textbf{l}}_a|=0$). The tangential spring
is activated when the particle comes into contact with particle $b$. The
tangential force is history-dependent. The contact friction force is
proportional to the tangential spring displacement, which is integrated over
the contact time as
\begin{equation}
  \label{eq:tangential-force}
  \ten{F}_{a}^{t^{n+1}} =
  -k_f \Delta \textit{\textbf{l}}_a^{\,n + 1} =
  -k_f \big[\big(\Delta {\textit{\textbf{l}}}_a^{\,n} \
  + \ten{v}_{ab}^{n + 1} \Delta t\big) \cdot \ten{t}_a^{n + 1} \big] \
  \ten{t}_a^{n + 1},
\end{equation}
where $\Delta t$ is the time step, $\ten{v}_{ab} = \ten{v}_{a} - \ten{v}_b$ is the
relative velocity of particle $a$ with respect to the contacting particle $b$,
and $k_f$ is the tangential spring stiffness coefficient. The tangential
spring stiffness is computed similar to the normal spring
stiffness\cite{golshan2023lethe}.  The tangential unit vector is computed by,
\begin{equation}
  \label{eq:tangential-vect}
  \ten{t}_a = \frac{\ten{v}_{ab} - (\ten{v}_{ab} \cdot \ten{n}) \ten{n}}{|\ten{v}_{ab} - (\ten{v}_{ab} \cdot \ten{n}) \ten{n}|}.
\end{equation}

@@

\begin{figure}[!htpb]
  \centering
  \begin{subfigure}{0.22\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{images/rfc_explantion_schematic/real_spherical_particles}
    \subcaption{A rigid spherical particle}%\label{fig:rings:ipst-nu-0.47-0}
  \end{subfigure}\hspace{15mm}%
  \begin{subfigure}{0.24\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{images/rfc_explantion_schematic/sph_sampled_spherical_particles}
    \subcaption{A rigid spherical particle sampled with dummy SPH particles}%\label{fig:rings:ipst-nu-0.47-1}
  \end{subfigure}
  \caption{}
\label{fig:real_particle_sph_sampling}
\end{figure}
To calculate the force exerted on the spherical particle by the surrounding
fluid, we employ a method involving the sampling of the spherical particle
using dummy Smoothed Particle Hydrodynamics (SPH) particles, depicted in
\cref{fig:real_particle_sph_sampling}. These SPH particles are evenly
distributed and remain stationary, moving in tandem with the velocity of the
rigid particle at any given location. To establish the pressure of these SPH
particles, we utilize the fixed ghost particle boundary technique outlined in
\cref{sec:boundary_conditions}.

>>>>

\documentclass[preprint,12pt]{elsarticle}
% \documentclass[preprint,12pt,draft]{elsarticle}
% \documentclass[draft,12pt]{elsarticle}

@@

  \title{{DEM}-{SPH} study of particles dispersion in fluid}
  \author[xxx]{Dinesh Adepu\corref{cor1}}
  \ead{d.dinesh@surrey.ac.uk}
  \author[xxx]{Chuan Yu Wu}
  % \ead{XXX}
\address[xxx]{School of Chemistry and Chemical Engineering, University of Surrey}

@@

\begin{abstract}
  Mixing powdered substances in tanks using a stirrer is a common occurrence
  across various industries. Typically, this phenomenon is studied numerically
  due to the complex nonlinear physics involved. In our current study, we have
  developed a solver combining Smoothed Particle Hydrodynamics (SPH) and
  Discrete Element Method (DEM) to investigate the mixing behavior of rigid
  circular particles within a tank under the influence of a stirrer operating
  at different speeds.  The SPH method governs the fluid phase, while the
  dynamics and interactions of particles are captured by the DEM. We achieve a
  fully resolved coupling between solid particles and fluid particles by
  discretizing solid particles into dummy SPH particles. Our validation
  process includes verifying the fluid solver's accuracy through a Poiseuille
  problem and validating the DEM solver by benchmarking interactions between
  two particles and particle-wall impacts.  The coupled model's is validated
  by simulating single-particle entry in a 2D steady tank and a cube settling
  in a tank, which is a 3D case. Following validation, we proceed with the
  mixing study. We conduct two simulations: one with spherical particles of
  uniform radius and another with particles having two different radius.  Our
  findings reveal that as for lower stirrer speeds, particles tend to
  aggregate initially and remain in the center. However, with further increase
  in stirrer speed, particles tend to accumulate near the tank corners due to
  the circulation induced in the fluid due to the stirrer motion. We use an
  open-source software, PySPH, adapting and integrating the coupled model to
  conduct the current study.
\end{abstract}

@@

Particle dispersion is a fundamental physical phenomenon encountered in
various industries where powder and fluid interactions are pivotal.  Petroleum
industry \cite{wang2023developments}, the transport of bodies in internal
systems \cite{Dai2021}, debris flow \cite{Qingyun2022}, the food processing
industry \cite{Karunasena2014}, and ice-sea modeling \cite{Mintu2018} are a
few areas to mention.  In many processes the mixing of powder with fluid using
a stirrer are commonplace \cite{li2022study}.  \Cref{fig:gloabl_body_frame_rb}
illustrates a sample particle dispersion problem studied in the current work.
While experimental investigation of such processes is often impractical due to
their highly nonlinear nature, numerical methods offer a viable alternative.
While these systems are studied with numerical methods, they are a part of
two-way coupling models. In numerical method approach, one can employ either a
mesh-based or meshless technique to study the current phenomena.

\begin{figure}[!htpb]
  \centering
  \includegraphics[width=0.4\textwidth]{images/hs_tank_fluid_with_particles}
  \caption{An illustrative figure showcasing the dispersion of particles within a free surface tank.}
  \label{intro:schematic}
\end{figure}

The combination of Discrete Element Method (DEM) for solid particle
interactions and Computational Fluid Dynamics (CFD) for fluid modeling
is a classic approach in handling particulate flow problem. In
mesh-based modeling, lattice Boltzmann method (LBM)
\cite{xiong2014lbm} and finite volume method (FVM) \cite{kloss2012models} are
utilized to handle the fluid dynamics. Similarly, in meshless schemes like
Smoothed Particle Hydrodynamics (SPH) \cite{peng2021fully},
Moving Particle Semi-implicit (MPS), and Particle Finite Element
Method (PFEM) \cite{li2019modeling, franci2020pfem}, handles the fluid
dynamics.

While handling the interaction between the fluid and solid particles, the
interaction can be categorized into a resolved and unresolved coupling.  In
unresolved coupling, simplified models are utilized to calculate the forces
exerted on solid particles within the fluid flow. In contrast, in the fully
resolved coupling approach, the fluid forces are directly computed on the
solid particles without the need for empirical drag force models.  Both
strategies are applicable in both meshless \cite{cleary2015prediction,
  trujillo2020smooth} and mesh-based \cite{van2008numerical, ma2022review}
CFD-DEM techniques.  In addition to the study mono-sized particle behaviour in
a fluid \cite{trujillo2020smooth}, research such as
\cite{brosh2014accelerating} and \cite{peng2021fully} explore wide particle
size distributions with CFD-DEM and SPH-DEM respectively.  Regarding
non-Newtonian fluid modelling in mesh-based schemes, works like
\cite{li2018dam} are notable and in meshless schemes, \cite{peng2021fully} is
notable. \citet{zhu2008discrete} and \citet{ma2022review} delve into
investigations on spherical and non-spherical particles respectively within
CFD-DEM frameworks.  Meshless CFD-DEM, though fruitful, encounters challenges
in handling problems with free surfaces and becomes computationally demanding
with a large number of solid particles, especially in fully resolved
coupling. In contrast, SPH offers advantages in such scenarios due to its
inherent capability in handling free surfaces and adeptness in modeling large
mesh deformation problems, facilitating the incorporation of novel physics.

The SPH method is a meshless numerical method originally proposed by
\citet{gingold1977smoothed} and \citet{lucy1977numerical} to model
astrophysics problems. It has been extensively applied to simulate problems
involving fluids, structural dynamics, fluid-structure interaction, granular
physics, non-Newtonian fluid flows \cite{peng2021fully} among other areas
\cite{monaghan2012smoothed}.  Various kind of particulate flows, involving,
simple spherical particles mixing using unresolved coupling
\cite{cleary2015prediction, markauskas2019coupled} and with debris flow
\cite{canelas2016sph} with fully resolved coupling approach were analysed
using SPH-DEM.  Handling particles of arbitrary shape in fluid flow is
addressed in various works, including those by \citet{peng2021fully,
  canelas2016sph, amicarelli2015smoothed}.  Several variants of SPH, such as
WCSPH \cite{peng2021fully, cleary2015prediction}, and ISPH
\cite{asai2021fluid}, are utilized for modeling fluid flow with particulate
flows. The coupling between fluid and rigid bodies in the fully resolved
category employs techniques such as the fixed ghost particle technique
\cite{canelas2016sph, asai2021fluid}, single layers of dummy SPH particles
\cite{peng2021fully}, and simple repulsive forces
\cite{monaghan2009sph}. Coupling between SPH fluid particles and solid
structures discretized as vertex edges is also proposed by
\citet{park2023new}. Particle settling with variable sizes is studied by
\citet{zou2022study}.  Elastic behavior of spherical particles and dispersion
studies are conducted by \citet{ng2021numerical}.  Interaction between rigid
bodies of arbitrary sizes is addressed using various techniques, such as those
discussed in papers by \citet{asai2021fluid}, \citet{peng2021fully},
\citet{canelas2016sph}.

In the current work we consider the particle dispersion within a large-scale
under stirrer mixing phenomenon. We employ Smoothed Particle Hydrodynamics
(SPH) to model fluid dynamics and the Discrete Element Method (DEM) to handle
the dynamics and contact interactions among spherical particles. The
interaction between solids and fluids is modelled using a fixed ghost particle
approach, ensuring a fully resolved simulation. We validate the fluid solver
independently using Poiseuille flow and the DEM solver separately using
fundamental benchmarks such as particle-particle and particle-wall
impacts. Subsequently, the fully resolved coupled SPH-DEM solver is validated
using three benchmarks: a single particle entering, two particles settling,
and a forced wedge entry into a steady hydrostatic tank. Upon confirming the
fidelity of the solver, we proceed with analyzing the mixing behavior of
various spherical particles influenced by a stirrer in a tank. We investigate
scenarios involving particles of different densities, stirrer velocities, and
variable sizes as part of our case study. The development work is conducted
using PySPH\cite{ramachandran2021pysph}, an open-source code available at
\url{https://github.com/pypr/pysph}.  We modify the current codebase to
integrate implementations of the Discrete Element Method (DEM) and the
Smoothed Particle Hydrodynamics (SPH) coupled with DEM model for our ongoing
research. To ensure reproducibility, we utilize the automan package
\cite{ramachandran2018automan} to automate all results generated in the
current manuscript, aiming for a reproducible research approach.

@@

\FloatBarrier%
\section{Fluid modeling}
\label{sec:fluid-modeling}
In the current work, we follow a weakly-compressible SPH approach to model the
fluid. The continuum in SPH is modeled using particles, which has physical
properties such as mass, velocity, and these particles interact based on the
governing equations using a Guassian-like kernel
\cite{monaghan-review:2005,morris1997modeling}.

\FloatBarrier%
\subsection{Fluid governing equations}
\label{sec:fluid--governing-equations}
The fluid motion is governed by the conservation of mass and the conservation
of linear momentum, which can be expressed by the following equations:
\begin{equation}
  \label{eq:ce}
  \frac{d \rho}{d t} = - \rho \; \frac{\partial u_i}{\partial x_i},
\end{equation}
\begin{equation}
  \label{eq:me}
  \frac{d u_i}{d t} = \frac{1}{\rho} \; \frac{\partial \sigma_{ij}}{\partial x_j}
  + g_i,
\end{equation}
where $\rho$ is the density, $u_i$ is the $i$\textsuperscript{th} component of
the velocity field, $x_j$ is the $j$\textsuperscript{th} component of the
position vector, $g_i$ is the component of body force per unit mass and
$\sigma_{ij}$ is stress tensor.

The stress tensor is split into pressure and viscous parts,
\begin{equation}
  \label{eq:fluid-stress-decomposition}
  \sigma_{ij} = - p \delta_{ij} + 2 \eta \frac{\partial u_i}{\partial x_j}
\end{equation}
where $\eta$ is the kinematic viscosity of the fluid, $p$ is the pressure,
$\delta_{ij}$ is the Kronecker delta function.

\FloatBarrier%
\subsection{Discretized fluid governing equations}
\label{sec:sph--governing-equations}
The governing equations involve function, derivative and divergence
operators. In SPH, these operators are approximated based on the positions,
mass and the kernel values of the discretized particles. Assume that the
domain is discretized into N particles with mass $m_i$ and volume
$V_i$, We have
\begin{equation}
  \label{eq:mass_repr}
  m_i = V_i \> \rho_i.
\end{equation}
Based on this discretization, the discrete function approximation is given as,
\begin{equation}
  \label{eq:discrete_form}
  A_i = \sum_{j \in \text{Neigh}(i)}\> \frac{m_j}{\rho_j} A_j\> W(\ten{x}_i - \ten{x}_j, h),
\end{equation}
where $A(\boldsymbol{x}_i) = A_i$ is the value of the field property of
particle i, similarly for $A_j$. Here, $\text{Neigh}(i)$ is the neighbours of
particle $i$.  A symmetric derivative approximation~\cite{Violeau16} is given
as,
\begin{equation}
  \nabla A(\ten{x}_i) = \rho_i \sum_{j} m_j \left(\frac{A_i}{\rho_i^2} + \frac{A_j}{\rho_j^2}\right) \nabla W_{ij}.
\end{equation}
A symmetric gradient operator ensures equal and opposite forces acting on the
particles.

@@

% \subsection{Discretized fluid governing equations}
The SPH discretization of the continuity
equation~\cref{eq:sph-discretization-continuity} is given as,
\begin{equation}
  \label{eq:sph-discretization-continuity}
  \frac{{d}\rho_a}{dt} = \sum_{b} \; \frac{m_b}{\rho_{b}} \;
  \rho_{a} \; {\ten{u}}_{ab} \; \cdot \nabla_{a} W_{ab},
\end{equation}

%
Similarly, the discretized momentum equation is written as,
\begin{multline}
  \label{eq:sph-momentum-fluid}
  \frac{{d}\ten{u}_{a}}{dt} = - \sum_{b} m_b
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg)
  \nabla_{a} W_{ab}
 \;+\;
  \sum_{b} m_b \frac{4 \eta \nabla W_{ab}\cdot
    \ten{r}_{ab}}{(\rho_a + \rho_b) (r_{ab}^2 + 0.01 h_{ab}^2)} \ten{u}_{ab}  \;+\;
  \ten{g}_{a},
\end{multline}
where $\ten{I}$ is the identity matrix, $\eta$ is the kinematic viscosity of the
fluid. The viscous term in \Cref{eq:me} is discretized according to the
formulation introduced in \cite{morris1997modeling}.

@@

The ghost particle approach proposed by \citet{Adami2012} is used to model the
boundaries. We use three layers of ghost particles to model the solid wall.
The properties of the solid wall are interpolated from the fluid particles.

When the viscous force is computed, the no slip boundary condition is used,
where the velocity on the boundary set as,
\begin{equation}
  \label{eq:no-slip-bc-u}
  \ten{u}_{\text{Ga}} = 2 \ten{u}_{\text{p}} - \ten{\hat{u}}_{\text{a}}.
\end{equation}
The projected velocity $\ten{\hat{u}}_{\text{a}}$ on the ghost particles is
computed using,
\begin{equation}
  \label{eq:v-ghost}
  \ten{\hat{u}}_a = \frac{\sum_b\ten{u}_b W_{ab}}{\sum_b W_{ab}},
\end{equation}
where $\ten{u}_b$, is the velocity of the fluid particle $b$ and $W_{ab}$ is the kernel
value between the fluid particle and the ghost particle.

@@

\FloatBarrier%
\subsection{Contact models}
\label{sec:dem}

@@

We resolve the contact among the spherical particles using the discrete
element method \cite{luding_dem_2008}. In the current work we have utilized a
non-linear contact force model. In DEM, the force acting on a particle $a$ due
to the interaction with the particle $b$ is resolved into a normal and
tangential component. The normal force component represents a repulsive force,
while the tangential component is used to model the friction between the
interacting particles.  The normal force ($\teng{F}_a^{n}$) on particle $a$
due to the interaction with the particles $b$ is given by a non-linear,
Hetzian model \cite{brilliantov1996model}, including damping is given as,
\begin{equation}
  \label{eq:contact-algorithm-normal}
  \ten{F}_a^n = k_n \delta_{n} \ten{n} - \eta_n (\ten{v}_{ab} \cdot \ten{n}) \ten{n}.
\end{equation}
Here, $\ten{v}_{ab} = \ten{v}_{a} - \ten{v}_b$ is the relative velocity of
particle $a$ with respect to the contacting particle $b$,
$\ten{n}$ normal unit vector passing from particle $b$ to $a$, and the overlap $\delta_{n}$ is computed
using the radius of particle $a$ ($R_a$) and particle $b$ ($R_b$) as
\begin{equation}
  \label{eq:cf-overlap}
  \delta_{n} = R_{a} + R_{b} - r_{ab},
\end{equation}
$k_n$ and $\eta_n$ are the normal spring stiffness and damping coefficient, which
are computed using the material properties of the bodies in contact and the
coefficient of restitution among the interacting bodies, given by:
\begin{equation}
  \label{eq:kf-stiffness}
  k_n = \frac{4}{3} \; E^{*} \; \sqrt{R^{*} \delta_n},
\end{equation}
\begin{equation}
  \label{eq:kf-stiffness}
  \eta_n = -2 \sqrt{\frac{5}{6}} \beta \; \sqrt{S_n m^*}.
\end{equation}
Where the other parameters are given as:
\begin{equation}
  \label{eq:kf-stiffness}
  \frac{1}{E^{*}} = \frac{1 -\nu_a^2}{E_a} + \frac{1 -\nu_b^2}{E_b}
\end{equation}
\begin{equation}
  \label{eq:kf-stiffness}
  \beta = \frac{\ln{e}}{\ln{e}^2 + \pi},
\end{equation}
here, $e$ is the coefficient of restitution.
\begin{equation}
  \label{eq:kf-stiffness}
  R^{*} = \frac{R_a R_b}{R_a + R_b}
\end{equation}
\begin{equation}
  \label{eq:kf-stiffness}
  m^{*} = \frac{m_a m_b}{m_a + m_b}
\end{equation}
\begin{equation}
  \label{eq:kf-stiffness}
  S_n = 2 E^{*} \sqrt{R^{*} \delta_n}
\end{equation}

\subsection{Tangential force computation}
\label{sec:tangential-force-computation}
To handle the frictional contact, we associate a tangential spring attached to
particle $a$ and particle $b$ to compute the tangential force, which initially has
a magnitude of zero ($|\Delta \textit{\textbf{l}}_a|=0$). The tangential spring
is activated when the particle comes into contact with particle $b$. The
tangential force is history-dependent. The contact friction force is
proportional to the tangential displacement, which is integrated over
the contact time as
\begin{equation}
  \label{eq:tangential-force}
  \ten{F}_{a}^{t^{n+1}} =
  -k_t \Delta \textit{\textbf{l}}_a^{\,n + 1} - \eta_t \ten{v}_t =
  -k_t \big[\big(\Delta {\textit{\textbf{l}}}_a^{\,n} \
  + \ten{v}_{ab}^{n + 1} \Delta t\big) \cdot \ten{t}_a^{n + 1} \big] \
  \ten{t}_a^{n + 1} - \eta_t \ten{v}_t,
\end{equation}
where $\Delta t$ is the time step,
and $k_t$ is the tangential spring stiffness coefficient. The tangential unit
vector is computed by,
\begin{equation}
  \label{eq:tangential-vect}
  \ten{t}_a = \frac{\ten{v}_{ab} - (\ten{v}_{ab} \cdot \ten{n}) \ten{n}}{|\ten{v}_{ab} - (\ten{v}_{ab} \cdot \ten{n}) \ten{n}|}.
\end{equation}
The tangential spring stiffness ($k_t$) and the tangential damping coefficient
$\eta_t$ are given as:
\begin{equation}
  \label{eq:kf-stiffness}
  k_t = 8 \; G^{*} \; \sqrt{R^{*} \delta_n},
\end{equation}
\begin{equation}
  \label{eq:kf-stiffness}
  \eta_t = -2 \sqrt{\frac{5}{6}} \beta \; \sqrt{S_t m^*},
\end{equation}
with $G^*$ and $S_t$ are given as,
\begin{equation}
  \label{eq:kf-stiffness}
  \frac{1}{G^{*}} = \frac{2 (2 - \nu_a) (1 + \nu_a)}{E_a} +  \frac{2 (2 - \nu_b) (1 + \nu_b)}{E_b},
\end{equation}
\begin{equation}
  \label{eq:kf-stiffness}
  S_t = 8 G^{*} \sqrt{R^{*} \delta_n},
\end{equation}
where, $G$ is the shear modulus of the particles, $\nu$ is the Poisson's ratio.

@@

\begin{figure}[!htpb]
  \centering
  \begin{subfigure}{0.22\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{images/rfc_explantion_schematic/real_spherical_particles}
    \subcaption{A rigid spherical particle}%\label{fig:rings:ipst-nu-0.47-0}
  \end{subfigure}\hspace{15mm}%
  \begin{subfigure}{0.24\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{images/rfc_explantion_schematic/sph_sampled_spherical_particles}
    \subcaption{A rigid spherical particle sampled with dummy SPH particles}%\label{fig:rings:ipst-nu-0.47-1}
  \end{subfigure}
  \caption{A rigid particle being discretized into several dummy SPH
    particles, to handle the interaction with surrounding fluid.}
\label{fig:real_particle_sph_sampling}
\end{figure}
To calculate the force exerted on the rigid particle due to the surrounding
fluid, we employ a method involving the sampling of the spherical particle
using dummy Smoothed Particle Hydrodynamics (SPH) particles, depicted in
\cref{fig:real_particle_sph_sampling}. These SPH particles are evenly
distributed and remain stationary, moving in tandem with the velocity of the
rigid particle at any given location.  They serve as SPH boundary particles
and contribute to the computation of fluid particle density and acceleration
according to \Cref{eq:sph-discretization-continuity,eq:sph-momentum-fluid}.
To establish the pressure of these SPH particles, we utilize the fixed ghost
particle boundary technique outlined in \cref{sec:boundary_conditions}.
